---
title: "Assignment 4"
author: "Cryo"
date: "`r Sys.Date()`"
output: html_document
---
Your assignment:

The following data was collected by the local hospital. 
This data set contains 5 variables based on the observation of 10 patients. 
In addition to the measurements of the patients checking in to the hospital that night, 
this data provides 
patients' histories regarding the frequency of their visits to the hospital in the last 12 months.
This data displays 
the measurement of blood pressure, 
first assessment by a general doctor (bad=1, good =0) titled "first," 
the second assessment by an external doctor (called "second"), 
and the last row provides the head of the emergency unit's decision regarding immediate care for the patient based on the values 0 or 1 (low = 0, high =1).
The names of your variables are as follows: 

"Freq","bloodp","first”, " second”, ”finaldecision”

The rows 

1.    "0.6","103","bad","low","low”
2.     "0.3","87","bad","low","high”
3.     "0.4","32","bad","high","low”
4.      "0.4","42","bad","high","high"
5.     "0.2","59","good","low","low”
6.      "0.6","109","good","low","high”
7.     "0.3","78","good","high","low”
8.      "0.4","205","good","high","high”
9.      "0.9","135",”NA","high","high"
10.    "0.2","176",”bad","high","high”

Here is a clarification hint:
Frequency <- c(0.6,0.3,0.4,......
BP <- c(103,87,32,42,.....
First <- c(1,1,1,.....
Second <- c(0,0,1,1,...
FinalDecision <- c(0,1,0,1,...

A. Your first assignment: Create a side-by-side boxplot (>boxplot(x, ...)) and and histogram (>hist(x,...)
In the following examples, I used different data containers for the boxplot and histogram visual display. 
1. Boxplot
I use the following data container (3,2,5,6,4,8,1,2,3,2,4)

A <- c(3, 2, 5, 6, 4, 8, 1, 2, 3, 2, 4)
boxplot(A)
Barplot.png  

2. Histogram I used following data container (2,4,5,7,12,14,16)
B <- c(2, 4, 5, 7, 12, 14, 16)
hist(B)

histogram.png  
B. Discuss the outcome of your results regarding patients BPs & MD’s Ratings.
C. Post your result in your blog and code on GitHub. 

https://rstue.blogspot.com/2026/02/module-4-programming-structure.html
```{r setup}

```

```{r}
#fix quotations in list & convert to vector
rows <- c("0.6","103","bad","low","low",
"0.3","87","bad","low","high",
"0.4","32","bad","high","low",
"0.4","42","bad","high","high",
"0.2","59","good","low","low",
"0.6","109","good","low","high",
"0.3","78","good","high","low",
"0.4","205","good","high","high",
"0.9","135","NA","high","high",
"0.2","176","bad","high","high")

#create matrix using ncol and byrow
m <- matrix(rows, ncol = 5, byrow = TRUE)
m

#create a dataframe from the matrix using as.data.frame
df <- as.data.frame(m)

#set the colnames according to instructions
colnames(df) <- c("Freq", "bloodp", "first", "second", "finaldecision")

#clean null values
df[df=="NA"] <- "bad"
df
#convert first two columns to numeric
df$Freq <- as.numeric(df$Freq)
df$bloodp <- as.numeric(df$bloodp)

#convert the remaining colums to boolean
df$first <- ifelse(df$first == "bad", 1, 0)
df$second <- ifelse(df$second == "high", 1, 0)
df$finaldecision <- ifelse(df$finaldecision == "high", 1, 0)

df
#create hist and boxplot using df columns
plot(df$bloodp, df$Freq)
hist(df$bloodp)
boxplot(df$bloodp)
boxplot(df$bloodp ~ df$finaldecision)
```
