---
title: "Final Project"
author: "VaShay Carpenter"
date: "`r Sys.Date()`"
output: html_document
---

## Install packages

```{r setup}
#install.packages("riem")
library(riem)
library(ggplot2)
library(dplyr)
library(lubridate)
```

## Locate a dataset that captivates your interest. Your dataset should contain at least 50 observations and 5 to 20 variables.

```{r}
# Pull riem data for Tampa International (TPA)
tpa_weather <- riem_measures(station = "TPA", date_start = "2026-01-01")

# Isolate relevant DF columns     
tpa_weather <- subset(tpa_weather, select = c(valid, alti, relh, tmpf, p01i, vsby))

# Remove Null Values
tpa_weather <- na.omit(tpa_weather)

# Rename columns
colnames(tpa_weather) <- c("Date", "Pressure", "Relative Humidity", "TemperatureF", "Precipitation", "Visibility")

# Show correlation between Atmospheric Pressure and Hourly Precipitation amounts
cor(tpa_weather$Pressure, tpa_weather$Precipitation, use = "complete.obs")

# Plot Pressure vs. Precipitation using a scatter plot with a regression line  
ggplot(tpa_weather, aes(x = Pressure, y = Precipitation)) +
  geom_point() +                          
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "Tampa Hourly Barometric Pressure vs Precipitation 2026", x = "Pressure", y = "Precipitation")
```

## R Markdown

```{r cars}
# Calculate daily averages and totals
tpa_daily <- tpa_weather %>%
  mutate(date_only = as.Date(Date)) %>%
  group_by(date_only) %>%
  summarise(
    avg_pressure = mean(Pressure, na.rm = TRUE),      # Altimeter (Pressure)
    avg_humidity = mean(`Relative Humidity`, na.rm = TRUE),      # Relative Humidity
    avg_temp     = mean(TemperatureF, na.rm = TRUE),      # Temperature (F)
    total_precip = sum(Precipitation, na.rm = TRUE),       # Total Rain per day
    avg_vis      = mean(Visibility, na.rm = TRUE),      # Visibility (Density proxy)
    .groups = "drop"
  )

# Recalculate correlation between Atmospheric Pressure and Daily Precipitation totals
cor(tpa_daily$avg_pressure, tpa_daily$total_precip, use = "complete.obs")

# Replot Average Daily Pressure vs. Total Daily Precipitation using a scatter plot with a regression line
ggplot(tpa_daily, aes(x = avg_pressure, y = total_precip)) +
  geom_point() +                          # Add scatter points
  geom_smooth(method = "lm", se = FALSE) + # Add linear regression line
  labs(title = "Tampa's Average Daily Pressure vs Total Daily Precipitation", x = "Pressure", y = "Precipitation")

```

## Including Plots

Create a visualization of your data and discuss your output. You can use any of the apps we covered in the class including: plot.ly, RStudio, Tableau or any app you like.

```{r pressure, echo=FALSE}

# Plot both the daily precipitation total as well as the average barometric pressure in series
ggplot(tpa_daily, aes(x = date_only)) +
  # BAR PLOT: Total Precip (Scaled up for visibility)
  geom_bar(aes(y = total_precip, fill = "Daily Precip"), 
           stat = "identity", alpha = 0.6) +
  # LINE PLOT: Average Pressure (Scaled down)
  geom_line(aes(y = avg_pressure - 30, color = "Barometric Pressure (Scaled down for visibility)"), size = 1.2, lwd = 0.2) +
  # FORMATTING
  scale_fill_manual(values = c("Daily Precip" = "skyblue")) +
  scale_color_manual(values = c("Barometric Pressure (Scaled down for visibility)" = "red")) +
  theme_minimal() +
  labs(
    title = "Pressure Trends vs. Precipitation Accumulation",
    subtitle = "TPA Station: Jan 2026 - May 2026",
    x = "Date",
    y = "Precipitation",
    fill = "",
    color = ""
  ) +
  theme(legend.position = "bottom")

# Applying daily plotting logic back to hourly dataset
ggplot(tpa_weather[50:110,], aes(x = Date)) +
  # BAR PLOT: Total Precip)
  geom_bar(aes(y = Precipitation, fill = "Daily Precip"), 
           stat = "identity") +
  # LINE PLOT: Average Pressure (Actual values)
  geom_line(aes(y = Pressure - 30, color = "Barometric Pressure (Scaled down for visibility)"), size = 1.2) +
  # FORMATTING
  scale_fill_manual(values = c("Daily Precip" = "skyblue")) +
  scale_color_manual(values = c("Barometric Pressure (Scaled down for visibility)" = "red")) +
  theme_minimal() +
  labs(
    title = "Pressure Trends vs. Precipitation Accumulation",
        subtitle = "TPA Station: Jan 2026 Precipitation Spike (Zoomed)",
    x = "Date",
    fill = "",
    color = ""
  ) +
  theme(legend.position = "bottom")

```

## Determine the accuracy and precision of the proposed methodology and update tpa_weather variable

```{r}
# Create variable markers to determine the onset of rain and 6 hours of atmospheric pressure readings preceding the onset of rain
tpa_weather <- tpa_weather %>%
  arrange(Date) %>%
  mutate(
    # 6-hour look-back slope
    pre_rain_slope = lag(Pressure, 1) - lag(Pressure, 7),
    # Filter significant declines in pressure only to reduce noise
    is_significant_signal = pre_rain_slope < -0.01,
    # Identification of rain onset
    rain_start = Precipitation > 0 & lag(Precipitation, default = 0) == 0,
    # Conditionally distinguish between significant signals
    signal_type = case_when(
      is_significant_signal & rain_start == TRUE ~ "True Positive (Hit)",
      is_significant_signal & rain_start == FALSE ~ "False Positive (Noise)",
      TRUE ~ "Ambient Noise"
  )
)
# Extract the starts for the Accuracy Calculation
tpa_leads <- tpa_weather %>% filter(rain_start == TRUE) 
# 1. Total Count of System 'Pings' (Signals)
total_signals_fired <- sum(tpa_weather$is_significant_signal, na.rm = TRUE)

# Total Count of Actual Rain Onsets
total_rain_events <- nrow(tpa_leads)

# Successes and Failures
successful_hits <- sum(tpa_weather$signal_type == "True Positive (Hit)", na.rm = TRUE)
false_alarms    <- sum(tpa_weather$signal_type == "False Positive (Noise)", na.rm = TRUE)

# Of all rain events, how many did we catch? 
prediction_rate <- (successful_hits / total_rain_events) * 100

# PRECISION: Of all the times the alarm went off, how often was it right? 
true_alarm_precision <- (successful_hits / total_signals_fired) * 100

# NOISE: Of all the times the alarm went off, how often was it WRONG? 
false_alarm_rate <- (false_alarms / total_signals_fired) * 100

# Print Results
print(paste0("CShell Hit Rate (Recall): ", round(prediction_rate, 2), "%"))
print(paste0("CShell Alarm Precision: ", round(true_alarm_precision, 2), "%"))
print(paste0("CShell False Alarm Rate (Noise): ", round(false_alarm_rate, 2), "%"))
```

## Plot tpa_weather distinguishing between True and False positives

```{r}
# Create a plot using the updated tpa_weather
  ggplot(tpa_weather[400:600,], aes(x = Date)) +
  geom_line(aes(y = Pressure - 30, color = "Pressure Trend"), lwd = 0.5) +
  
  # Highlight segments where the signal type indicated a true positive predictor
  geom_point(data = filter(tpa_weather[400:600,], signal_type == "True Positive (Hit)"), 
             aes(y = Pressure - 30, color = "True Positive (Hit)"), size = 2, alpha = 0.8) +
  geom_point(data = filter(tpa_weather[400:600,], signal_type == "False Positive (Noise)"), 
             aes(y = Pressure - 30, color = "False Positive (Noise)"), size = 1, alpha = 0.4) +
  
  # Rain Bars (Set color to NA to prevent red outlines)
  geom_bar(aes(y = Precipitation, fill = "Rainfall"), stat = "identity", color = NA) +
  # Formatting
  scale_color_manual(values = c(
    "Pressure Trend" = "grey40",
    "True Positive (Hit)" = "forestgreen",
    "False Positive (Noise)" = "firebrick"
  ))+
    #Fix fill color mismatch
    scale_fill_manual(values = c("Rainfall" = "skyblue")) +
  labs(title = "CShell Tactical Telemetry: Lead-Indicator Analysis",
       subtitle = "Red Indicators = 6-Hour Barometric Decline (Predictive Phase)",
       y = "Inches / Pressure Deviation (Baseline 30.00)") +
  theme_minimal()

# Summarizing to Daily 
tpa_daily_analysis <- tpa_weather %>%
  mutate(date_only = as.Date(Date)) %>%
  group_by(date_only) %>%
  summarise(
    avg_pressure = mean(Pressure, na.rm = TRUE),
    total_precip = sum(Precipitation, na.rm = TRUE),
    # Did we have a successful prediction today?
    had_hit = any(signal_type == "True Positive (Hit)"),
    had_noise = any(signal_type == "False Positive (Noise)"),
    .groups = "drop"
  )

ggplot(tpa_daily_analysis, aes(x = date_only)) +
  # 1. Barometric Trend Line
  geom_line(aes(y = avg_pressure - 30), color = "grey70", size = 0.8) +
  
  # 2. Daily Rainfall Bars
  geom_bar(aes(y = total_precip, fill = "Daily Rainfall"), stat = "identity") +
  
  # 3. Successful Hits (Green)
  geom_point(data = filter(tpa_daily_analysis, had_hit), 
             aes(y = avg_pressure - 30, color = "Successful Prediction"), size = 2) +
             
  # 4. False Positives (Red/Orange)
  geom_point(data = filter(tpa_daily_analysis, had_noise & !had_hit), 
             aes(y = avg_pressure - 30, color = "False Positive"), size = 1, alpha = 0.5) +

  scale_color_manual(values = c("Successful Prediction" = "forestgreen", "False Positive" = "firebrick")) +
  scale_fill_manual(values = c("Daily Rainfall" = "skyblue")) +
  labs(title = "CShell Semester Analysis: Predictive Success vs. Noise",
       subtitle = "Tampa International (TPA): Jan - May 2026",
       y = "Inches / Pressure Deviation") +
  theme_minimal()

```

## Function to repeat Model Logic

```{r}
analyze_lead_indicator <- function(station_id, start_date = "2026-01-01") {
  
  # 1. Pull and Initial Clean
  raw_data <- riem_measures(station = station_id, date_start = start_date)
  
  clean_data <- raw_data %>%
    subset(select = c(valid, alti, p01i)) %>%
    na.omit() %>%
    arrange(valid) %>%
    mutate(
      # 6-hour look-back slope
    pre_rain_slope = lag(alti, 1) - lag(alti, 7),
    # Filter significant declines in pressure only to reduce noise
    is_significant_signal = pre_rain_slope < -0.01,
    # Identification of rain onset
    rain_start = p01i > 0 & lag(p01i, default = 0) == 0,
    # Conditionally distinguish between significant signals
    signal_type = case_when(
      is_significant_signal & rain_start == TRUE ~ "True Positive (Hit)",
      is_significant_signal & rain_start == FALSE ~ "False Positive (Noise)",
      TRUE ~ "Ambient Noise"
    )
  )
  
  # Extract the starts for the Accuracy Calculation
tpa_leads <- clean_data %>% filter(rain_start == TRUE) 
# Calculate Prediction Accuracy using updated significance variables
successful_alerts <- sum(clean_data$signal_type == "True Positive (Hit)", na.rm = TRUE)
false_alerts <- sum(clean_data$signal_type == "False Positive (Noise)", na.rm = TRUE)

total_events <- nrow(tpa_leads)
# Successful prediction rate significantly diminished testing pre_rain_slope < -0.02 changed to -0.01
prediction_rate <- (successful_predictions / total_events) * 100
# Rain events that happened without a pressure drop.
# 1. How many total "Alerts" did the system fire?
total_signals_fired <- sum(clean_data$is_significant_signal, na.rm = TRUE)

# 2. How many of those alerts were WRONG (No rain followed)?
false_alerts <- sum(clean_data$signal_type == "False Positive (Noise)", na.rm = TRUE)

# 3. Noise Rate = (False Alarms / Total Alarms Fired)
true_alarm_rate <- (successful_alerts / total_signals_fired)* 100
false_alarm_rate <- (false_alerts / total_signals_fired) * 100
# Print Prediction Accuracy using 6 hour look-back time
print(paste0(station_id," True Alarm Rate: ", round(true_alarm_rate, 2), "%"))
print(paste0(station_id, " False Alarm Rate: ", round(false_alarm_rate, 2), "%"))
  
  # 2. Calculate Prediction Accuracy
  starts <- clean_data %>% filter(rain_start == TRUE)
  successes <- sum(starts$pre_rain_slope < 0, na.rm = TRUE)
  total_events <- nrow(starts)
  
  accuracy <- (successes / total_events) * 100
  
  # 3. Output results as a list
  return(list(
    true_accuracy = round(true_alarm_rate, 2),
    false_accuracy = round(false_alarm_rate, 2),
    data = clean_data,
    total_events = total_events
  ))
}
```

# Test Precision on other locations

```{r}
jfk_res <- analyze_lead_indicator("KJFK")
mia_res <- analyze_lead_indicator("MIA")
jax_res <- analyze_lead_indicator("JAX")
tpa_res <- analyze_lead_indicator(("TPA"))
```

Conclusion: While the CShell model successfully identified 70% of precipitation onsets, its 3% Precision and over 95% False Alarm Rate renders it tactically unviable as a standalone alert. This project demonstrates the critical need for multivariate sensor fusion in meteorological telemetry.
