---
title: 'Module #6 Doing Math'
author: "Cryo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

```
Answer the following questions and post your answer on your blog:

1. Consider A=matrix(c(2,0,1,3), ncol=2) and B=matrix(c(5,2,4,-1), ncol=2).
    a) Find A + B
    b) Find A - B

2. Using the diag() function to:
    build a matrix of size 4 with: 
    the following values in the diagonal 4,1,2,3.

3. Generate the following matrix:

  ## [,1] [,2] [,3] [,4] [,5]
  ## [1,] 3 1 1 1 1
  ## [2,] 2 3 0 0 0
  ## [3,] 2 0 3 0 0
  ## [4,] 2 0 0 3 0
  ## [5,] 2 0 0 0 3
  
  Hint: You can use the command diag() to build it.

```{r 1}
A <- matrix(c(2,0,1,3), ncol=2)
A
B <- matrix(c(5,2,4,-1), ncol=2)
B

sumAB <- A + B
sumAB
diffAB <- A - B
diffAB


```

2. Using the diag() function to build a matrix of size 4 with the following values in the diagonal 4,1,2,3.

```{r 2}
numbertwo <- diag(c(4,1,2,3))
numbertwo
```
3. Generate the following matrix:

## [,1] [,2] [,3] [,4] [,5]
## [1,] 3 1 1 1 1
## [2,] 2 3 0 0 0
## [3,] 2 0 3 0 0
## [4,] 2 0 0 3 0
## [5,] 2 0 0 0 3
Hint: You can use the command diag() to build it.

```{r 3}
# Step 1: Create a 5x5 matrix of zeros
mat <- matrix(0, nrow = 5, ncol = 5)

# Step 2: Set the diagonal to 3
diag(mat) <- 3
#Step 3: Set first row to 1 from 2:5
mat[1, 2:5] <- 1
# Step 4: Set first column to 2 from 2:5
mat[2:5, 1]<- 2

mat
```
