---
title: "Module 11 Debug"
author: "VaShay Carpenter"
date: "`r Sys.Date()`"
output: html_document
---

Below is a function intended to flag rows of a numeric matrix x that are outliers in every column according to the Tukey rule. The function contains a deliberate bug. A helper function for detecting Tukey outliers is provided first so that you can focus on debugging the logical error in the main function.

```{r setup, include=FALSE}
turkey.outlier <- function(x, k = 1.5) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr <- q3 - q1
  x < (q1 - k * iqr) | x > (q3 + k * iqr)
}

turkey_multiple <- function(x) {
  outliers <- array(TRUE, dim = dim(x))
  for (j in 1:ncol(x)) {
    outliers[, j] <- outliers[, j] && turkey.outlier(x[, j])
  }
  outlier.vec <- vector("logical", length = nrow(x))
  for (i in 1:nrow(x)) {
    outlier.vec[i] <- all(outliers[i, ])
  }
  return(outlier.vec)
}
```

Reproduce the Error
In R, create a test matrix and run the function:
```{r cars}
set.seed(123)
test_mat <- matrix(rnorm(50), nrow = 10)
turkey_multiple(test_mat)

#Capture the exact error message or warning message you see.
```

Diagnose the Bug
Explain why using && inside the loop is incorrect. Recall that && evaluates only the first element of a logical vector, whereas this task requires element-wise comparison across all rows.

{}

Fix the Code
Edit the function so that the logical operation is applied element-wise. Replace the buggy line with:

outliers[, j] <- outliers[, j] & tukey.outlier(x[, j])

```{r pressure, echo=FALSE}
plot(pressure)
```

Validate Your Fix
Re-run your corrected function on test_mat and verify that it returns a logical vector of length 10 without error:

```{r}
corrected_tukey <- function(x) {
  outliers <- array(TRUE, dim = dim(x))
  for (j in seq_len(ncol(x))) {
    outliers[, j] <- outliers[, j] & tukey.outlier(x[, j])
  }
  outlier.vec <- logical(nrow(x))
  for (i in seq_len(nrow(x))) {
    outlier.vec[i] <- all(outliers[i, ])
  }
  outlier.vec
}

corrected_tukey(test_mat)

```
Defensive Enhancements (Optional)
Add checks at the top of your function to:
Ensure that x is a matrix.
Ensure that x is numeric.
Provide informative errors if assumptions are not met.
Example:

```{}
corrected_tukey <- function(x) {
  if (!is.matrix(x)) {
    stop("x must be a matrix.")
  }
  if (!is.numeric(x)) {
    stop("x must be a numeric matrix.")
  }

  outliers <- array(TRUE, dim = dim(x))
  for (j in seq_len(ncol(x))) {
    outliers[, j] <- outliers[, j] & tukey.outlier(x[, j])
  }

  outlier.vec <- logical(nrow(x))
  for (i in seq_len(nrow(x))) {
    outlier.vec[i] <- all(outliers[i, ])
  }

  outlier.vec
}
```
Document Your Debugging Workflow
On your blog, include:
The original error or warning message.

Your diagnosis of the bug and why the code was wrong.

The corrected code and its successful output.

Any additional defensive checks you added.