---
title: "CShell-Tactical-Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CShell-Tactical-Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Operational Overview
CShell is a biological optimization engine. Its primary objective is to maintain user homeostasis by removing the cognitive overhead of gear selection. It applies a hierarchical logic table to live environmental telemetry. The CShell package aims to simplify shell selection for even the most indecisive of mollusks. 

# Core Usage
To generate a tactical report for a specific location:
```{r}
library(CShell)
city_outfit("Largo")
```
# Versatile implementation
CShell supports both human-readable city inputs and raw geospatial coordinates for remote field operations.
```{r}
# Deployment by City Node
report_city <- city_outfit("Boston")

# Deployment by Precise Coordinates
report_coord <- coords_outfit(lat = 27.91, long = -82.79)

# Technical Verification
# The output of CShell functions is a specialized S3 object of class `outfit`. 

# Verification of S3 Class Usage
class(report_city)

# This allows for clean reporting while maintaining raw data accessibility.

# Access National Weather Service report link inside the object
report_city$Weather$source_url

# Access Suggested Outfit for Tactical Report
report_city$Outfit
```
**Internal Architecture & S3 Classes**
CShell utilizes S3 Classes to manage data output. When a function is called, it returns an object of class outfit. This allows for standardized data handling while providing a high-visibility telemetry report in the console.

**Future Proofing: API Integration**
The current version of CShell utilizes a web-scraping engine (scrape_nws) to pull data from the National Weather Service.

**For Contributors:**
The engine.R file is architected to be modular. While scraping is functional for the current module, future iterations should transition to a REST API (e.g., OpenWeather). To contribute:

Replace the scrape_nws logic with an API call.

Ensure the output dataframe maintains the existing column schema (temp, humidity, precip).

The apply_wardrobe_logic will automatically handle the new data source without further modification.

Add forecasts for future dates to create a weekly wardrobe report