---
output: github_document
---

# CShell

<!-- badges: start -->
<!-- badges: end -->

**CShell** is a tactical weather telemetry and logistical engine built in R. It automates the extraction of National Weather Service (NWS) data and processes it through an internal logic-gate system to provide real-time wardrobe and equipment recommendations.

## Technical Identity
*   **Domain:** Data Analytics & Logistics Automation
*   **Logic Engine:** Symbolic logic circuits (AND/OR gates) mapped to environmental variables.
*   **Architecture:** Modular R package with internalized logic tables for data sovereignty.

## Installation

You can install the development version of CShell from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("cryo-cell/r-programming-assignments", subdir = "CShell")
```

## Example Code

This is a basic example which shows you how to solve use the CShell package:

```{r example}
library(CShell)
## basic example code

# Execute the core engine for a specific node
weather <- scrape_nws("Largo", 27.91, -82.79)

# Generate wardrobe requirements based on internal logic gates
outfit <- city_outfit("Largo")
print(outfit)
```
Detailed Documentation
For a full tactical breakdown of the logic engine and data-raw processing, please refer to the package vignette:

```{r eval=FALSE}
vignette("CShell-Tactical-Guide", package = "CShell")
```